% ============================================================
% Verification Tables + Prose (Termination + Safety)
% F^{V,Delta,Sigma}_{Tendermint}  Delay-Attack Model
% ============================================================
%
% Required in preamble:
%   \usepackage{booktabs}
%   \usepackage{amsmath,amssymb}
%
% Insert in body:
%   \input{verification_tables}
% ============================================================

% ==================================================================
% 1) TERMINATION
% ==================================================================
\subsection{Termination Verification}

To verify termination of $\mathcal{F}^{V,\Delta,\Sigma}_{\mathrm{Tendermint}}$
under delay attacks, we use the Apalache model checker in simulation mode
(\texttt{apalache-mc simulate}) to construct explicit termination witnesses
from the TLA\textsuperscript{+} specification.
We fix the validator set $V=\{v_1,v_2,v_3,v_4\}$ with fault threshold $f=1$,
network delay bound $\Delta=1$, protocol attack tolerance $\Sigma=1$,
and block height $H_{\max}=1$ to cover cross-height state transitions.

Adversarial behaviour is governed by the transition relation
\texttt{NextBoundedAttack}.
Subject to $\texttt{attackCount}<f$, the adversary may inject a delay
$\sigma_p = 2\Sigma+1$ into any honest validator $p$, activating the guard
$\mathrm{DelayAttack}(p)\colon (\delta+\sigma_p > 2\Delta)\vee(\sigma_p > 2\Sigma)$
and forcing $p$ to execute the degraded branch in each pipeline step
(casting a nil vote or advancing the round without decision).
The parameter $R_{\max}$ bounds the number of rounds over which the adversary
may mount such an attack, and serves as the primary axis of the sweep.

We verify against the negated invariant
$\texttt{NotYetTerminated}\triangleq
\lnot(\forall p\in\mathrm{Honest}:\,\mathrm{decision}_p[0]\neq\mathrm{nil})$,
which asserts that not all honest validators have yet decided at height~$0$.
Whenever Apalache finds a counterexample to this invariant along a sampled
execution trace, it witnesses a run in which every honest validator reaches
a decision, thereby establishing termination for the given parameters.
Exit code RC$\,=\,12$ (Apalache's counterexample-found status) corresponds
precisely to this outcome.

Table~\ref{tab:termination} reports results for
$R_{\max}\in\{1,2,3,5,8,10\}$.
A termination witness was found in every case, confirming that the protocol
terminates under delay-attack adversaries with a round budget of up to
$R_{\max}=10$.

We note that verification time does not increase monotonically with $R_{\max}$:
for instance, TR-04 ($R_{\max}=5$, $117.1\,$s) took longer than
TR-05 ($R_{\max}=8$, $81.4\,$s).
This is expected behaviour of the simulation mode.
Unlike exhaustive model checking, \texttt{apalache-mc simulate} samples
execution traces at random and terminates as soon as one trace violates
the negated invariant \texttt{NotYetTerminated}---i.e., as soon as a
termination witness is found.
Consequently, wall-clock time reflects when the \emph{first} witness happens
to be sampled, which depends on the random seed and the structure of the
state space at that particular $R_{\max}$, rather than on the size of the
full reachable state space.
A larger $R_{\max}$ does not necessarily require more time if a short
witness trace exists and is discovered early in the sampling process.

\begin{table}[ht]
\centering
\caption{%
  Termination verification results for
  $\mathcal{F}^{V,\Delta,\Sigma}_{\mathrm{Tendermint}}$
  under bounded delay attack
  ($|V|=4$, $f=1$, $\Delta=\Sigma=1$, $H_{\max}=1$,
  \texttt{length}$=100$, \texttt{max-run}$=60$).
  RC$\,=\,12$ indicates that Apalache found a termination witness.
}
\label{tab:termination}
\small
\begin{tabular}{@{}ccrrc@{}}
\toprule
\textbf{Run} &
$R_{\max}$ &
\textbf{length} &
\textbf{Time\,(s)} &
\textbf{Result} \\
\midrule
TR-01 &  1 & 100 &    31.3 & PASS \\
TR-02 &  2 & 100 &    25.4 & PASS \\
TR-03 &  3 & 100 &    34.0 & PASS \\
TR-04 &  5 & 100 &   117.1 & PASS \\
TR-05 &  8 & 100 &    81.4 & PASS \\
TR-06 & 10 & 100 &  1097.1 & PASS \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{How to read Table~\ref{tab:termination}.}
Each column has a concrete meaning:
\begin{itemize}
  \item \textbf{Run}: experiment identifier (TR-01, TR-02, \dots) used for cross-reference in text and logs.
  \item $\mathbf{R_{\max}}$: maximum round bound used in that run (the attack/round budget parameter being swept).
  \item \textbf{length}: Apalache simulation trace bound (maximum transition steps per sampled execution).
  \item \textbf{Time (s)}: wall-clock runtime in seconds until the run terminates (typically when a witness is found or budget is exhausted).
  \item \textbf{Result}: outcome summary; \textbf{PASS} means a termination witness was found (RC$=12$ for this setup).
\end{itemize}

% ==================================================================
% 2) SAFETY
% ==================================================================
\subsection{Safety Verification}

To verify safety of $\mathcal{F}^{V,\Delta,\Sigma}_{\mathrm{Tendermint}}$
under delay attacks, we use Apalache in bounded checking mode
(\texttt{apalache-mc check}) and evaluate the invariant
$\texttt{Agreement}$, requiring that no two initially honest validators
decide different values at the same height.
We fix the validator set $V=\{v_1,v_2,v_3,v_4\}$ with fault threshold $f=1$,
network delay bound $\Delta=1$, protocol attack tolerance $\Sigma=1$,
and block height bound $H_{\max}=1$.

The checker outcome is interpreted as follows:
RC$\,=\,0$ (NoError) means no Agreement counterexample exists within the
bounded horizon (here \texttt{length}$=5$), while RC$\,=\,12$ would indicate
that a violating trace was found.
Table~\ref{tab:safety} shows RC$=0$ for every tested $R_{\max}$,
so no safety violation is observed in this configuration.

Similar to other bounded symbolic checks, runtime is not strictly monotonic
in $R_{\max}$: solver effort depends on the structure of enabled/disabled
transitions and pruning opportunities, not only on the numeric round bound.

\begin{table}[ht]
\centering
\caption{%
  Safety verification results for
  $\mathcal{F}^{V,\Delta,\Sigma}_{\mathrm{Tendermint}}$
  under bounded delay attack
  ($|V|=4$, $f=1$, $\Delta=\Sigma=1$, $H_{\max}=1$,
  \texttt{length}$=5$).
  RC$\,=\,0$ indicates that no Agreement counterexample was found
  within the bounded computation.
}
\label{tab:safety}
\small
\begin{tabular}{@{}ccrrc@{}}
\toprule
\textbf{Run} &
$R_{\max}$ &
\textbf{length} &
\textbf{Time\,(s)} &
\textbf{Result} \\
\midrule
SR-01 &  1 & 5 & 61.6 & PASS \\
SR-02 &  2 & 5 & 60.6 & PASS \\
SR-03 &  3 & 5 & 57.7 & PASS \\
SR-04 &  5 & 5 & 48.1 & PASS \\
SR-05 &  8 & 5 & 58.3 & PASS \\
SR-06 & 10 & 5 & 44.9 & PASS \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{How to read Table~\ref{tab:safety}.}
The columns are interpreted as follows:
\begin{itemize}
  \item \textbf{Run}: experiment identifier (SR-01, SR-02, \dots).
  \item $\mathbf{R_{\max}}$: round bound used in this bounded safety check.
  \item \textbf{length}: checking horizon (maximum behavior length explored by bounded model checking).
  \item \textbf{Time (s)}: wall-clock verification time in seconds.
  \item \textbf{Result}: outcome summary; \textbf{PASS} means no Agreement counterexample was found within the bound (RC$=0$).
\end{itemize}
